### 2025-05-29 choosing_ordinal_effect_measures.tex
Rename + start effect measures draft

\documentclass[
  11pt,
  fleqn
]{article}
% [fleqn] left-aligns all equations

\input{preamble/preamble}

% For writing/planning purposes: show paragraph level in TOC
\setcounter{secnumdepth}{3}

\addbibresource{bibliography.bib}
\graphicspath{{fig/}}

\title{Choosing ordinal effect measures for clinical trials}
\author{Leon Di Stefano + Ordinal
Working Group}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\newpage

\section{Introduction}

\end{document}