### 2025-05-29 constructing_ordinal_outcomes.tex

\documentclass[
  11pt,
  fleqn
]{article}
% [fleqn] left-aligns all equations

\input{preamble/preamble}

% For writing/planning purposes: show paragraph level in TOC
\setcounter{secnumdepth}{3}

\addbibresource{references.bib}
\graphicspath{{fig/}}

\title{Constructing ordinal outcomes for clinical trials}
\author{Leon Di Stefano + Ordinal
Working Group}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\newpage

\section{Introduction}

Ordinal outcomes have become increasingly prominent in clinical
resarch, particularly following the COVID-19 pandemic, when ordinal
scales developed by the WHO and NIAID were used in prominent trials
of drugs repurposed to treat the disease \todo{cite}. They have also
been a popular choice in adaptive platform trials \todo{cite}.

Some advantages of ordinal outcomes are that they can
\begin{enumerate}
  \item Provide more information that binary outcomes, including
    composite outcomes, making study designs more sensitive and precise, and
    therefore more efficient and ethical

  \item Allow for the incorporation of intercurrent events into one's primary
    outcome, avoiding complexities of alternative estimand strategies like
    principal stratum estimands (particular for truncating events
    like death) while
    yielding more statistical efficiency than traditional binary
    composite outcomes

  \item Allow for the easy incorporation of safety events into a
    primary outcome, accounting for many different aspects
    of benefit/harm; at the same time, ordinal outcomes can be
    analyzed without a
    fully-specified utility function, which may require more
    extensive development and validation or may vary widely across patients.

  \item Allow for the combination of discrete and continous outcome
    measures (for
      example a numerical quality of life scale with death as the
      overriding worst
    outcome) and allow for the combination of longitudinal and timepoint or
    ``state'' outcomes in a single measure.

\end{enumerate}

The goal of this manuscript is to serve as a guide for clinicians and trialists
in developing ordinal outcomes for their studies. We consider both the outcome
itself (the data recorded on each patient) and effect measures or ``endpoints''
(the metric used to compare distributions of the outcome across patient
groups).

We want to distinguish \emph{choosing} or \emph{constructing} an outcome/effect
measure from \emph{validating} the outcome/effect measure using external data.
This paper is about the first task, choosing (or constructing) the outcome and
effect measure. In practice choosing an outcome and effect measure and
validating that outcome and effect measure in pre-trial data will be an
iterative process. Our focus in this paper is on the ``positive'' side of the
process--namely coming up with, and tweaking, proposed outcomes and effect
measures--rather than the ``negative'' side, namely testing or checking
proposals with data to assess whether they have desired properties.

A final goal of this paper is to situate the literature around ``generalized
pairwise comparisons of prioritized outcomes''
\citep{buyseGeneralizedPairwiseComparisons2022}--including win ratio
\citep{pocockWinRatioNew2012} and Desirability of Outcome Ranking (DOOR)
methods \citep{evansDesirabilityOutcomeRanking2015,
ongUnlockingDOORHow2023}--within the broader context of ordinal outcomes and
endpoints.

\subsection{Desiderata for (ordinal) outcomes: validity and efficiency}

We begin with some general desiderata for outcomes and effect measures in
clinical studies.

We want outcomes that
\begin{enumerate}
  \item are \textbf{valid} in that they reflect patient benefit and harm:
    better/worse outcomes are judged respectively better/worse by patients, are
    objectively respectively better/worse for patients, or are reliable
    surrogates/predictors of outcomes that are respectively better/worse for
    patients
  \item are \textbf{efficient} in the sense of capturing
    sufficient information to
    discriminate among better- and worse-off patients without large
    sample sizes.
\end{enumerate}
(For the purposes of this article, set aside ease of collection, low
missingess, objective ascertainability, etc.)

An example of an outcome which is valid but inefficient might be
death as a binary outcome in a context where deaths
are rare.

An example of an outcome which is efficient but which may be invalid
might be continuously-measured concentrations of a biomarker which is a poor
surrogate for disease progression.

One point of view is that one would ``ideally'' analyze ordinal outcomes by
reporting the associated expected utilities \todo{cite Berry talk; example of
utility-weighted Rankin scale} and that since existing ordinal methods supply
such utilities implicitly why not just supply them explicitly. A counterpoint
is that the methods do not, in fact, supply utilities implicitly.
Another is that ordinal outcomes let one ``hedge'' about precise
utillities and this is OK.

\subsection{Constructing versus validating}

\section{Ways to combine ordinal outcomes}

\subsection{Sequential,
hierarchical, \& parallel composition}

Term in the literature: ``hierarchical composite endpoints''
\citep{gasparyanDesignAnalysisStudies2022}.

Three broad ways to combine ordinal outcomes:
\begin{enumerate}
  \item \textbf{Sequential composition} Tie-breaking by $B$
    \emph{within either highest or lowest level of $A$}. Also could
    call this ``overriding by $A$''. A special case of this is
    \emph{overriding events} when $A$ is binary.
  \item \textbf{Hierarchical composition} Complete tie-breaking of $A$ by
    $B$ (``prioritized outcomes''; lexicographic order) within
    \emph{every} level of $A$.

    This is what win methods do.
  \item \textbf{Parallel composition} This is where levels of $A$ are
    combined with levels of $B$ ``in parallel''; high levels of $A$
    are combined with high levels of $B$ and low levels of $A$ with
    low levels of $B$.

    This is not so precisely determined as the other two. We need to
    decided \emph{how} to aggregate.
\end{enumerate}

Suggested notation:
\begin{itemize}
  \item "$\rightarrow$" and "$\leftarrow$" to denote sequential
    composition/overriding,
  \item "$/$" to denote hierarchical composition/complete tiebreaking
    (note that this is an asymmetric operation), and
  \item "$|$" to denote parallel composition
\end{itemize}

Examples:

\begin{itemize}
  \item \textbf{Sequential composition} Outcome from PASSPORT,
    depicted in Figure
    \ref{fig:passport_outcome}, is given by
    \[ \text{day 28 status} \leftarrow \text{max pSOFA} \leftarrow \text{days
      hospitalized}
    \]
  \item
    \textbf{Hierarchical composition} ``Angina Symptom Score'' outcome from
    \citet{rajkumarPlaceboControlledTrialPercutaneous2023}, depicted in Figure
    \ref{fig:rajkumar_et_al_outcome}, is given by \[
      \text{adverse clinical event} \leftarrow ((\text{units of
    antianginal medication}) / (\text{episodes of angina})) \]

  \item \textbf{Parallel composition} BANDICOOT outcome
    \citep{walkerCodesigningNovelOrdinal2025}, which is
    \[
      (\text{death, graft loss, or relapse}) \leftarrow (\text{viraemia} \;|\;
      \text{CD4+ count} \;|\; \text{organ support})
    \]
\end{itemize}

Aside: \citet{wittkowskiCombiningSeveralOrdinal2004} consider combining
ordinal variables into partial orders. This makes more sense in the
generalized pairwise comparisons framework.

\begin{figure}
  \includegraphics[width=7in]{passport_ordinal_outcome_sequential.pdf}
  % \includegraphics[width=1in]{passport_ordinal_outcome_colorbar.png}
  \caption{Ordinal outcome used in the planned PASSPORT adaptive
    platform trial. This illustrates sequential composition/overriding:
    of days hospitalized by maximum pSOFA while hospitalized, and of
  both of these by day 28 status.}
  \label{fig:passport_outcome}
\end{figure}

\begin{figure}
  \includegraphics[width=6in]{rajkumar_et_al_outcome.png}
  \caption{``Angina Symptom Score'' ordinal outocome develped in
    \citep{rajkumarPlaceboControlledTrialPercutaneous2023}. This illustrates
    (1) complete tie-breaking (of units of antianginal medication by
    episodes of angina) and (2) overriding (of angina episodes and
  angina medication by adverse clinical events)}
  \label{fig:rajkumar_et_al_outcome}
\end{figure}

\begin{figure}
  \includegraphics[width=7in]{parallel_composition_bandicoot.pdf}
  % \includegraphics[width=1in]{passport_ordinal_outcome_colorbar.png}
  \caption{Ordinal outcome used in the planned BANDICOOT adaptive
    platform trial \citep{walkerCodesigningNovelOrdinal2025}.
  Illustrates \textbf{parallel composition} of component outcomes.}
  \label{fig:parallel_bandicoot}
\end{figure}

\subsection{Converting back to numeric scores}

Treat each input factor as
numeric then sum them up. I believe \citet{ongUnlockingDOORHow2023}
has examples?

This doesn't respect the ordinal nature of the variables. That is, it
implicitly uses information about the \emph{differences} between levels.

\subsection{Pairwise comparison methods (win, DOOR) implicitly construct ordinal
outcomes}

Terms to define/distinguish:
\begin{itemize}
  \item Desirability of Outcome Ranking (outcome)
  \item Hierarchical composite endpoint (outcome)
  \item Prioritized outcomes (outcome)
  \item Win methods (win probability, win odds, win ratio) (effect
    measure; estimation methods)
  \item Generalized pairwise comparisons (effect measure)
\end{itemize}

These approaches implicitly \textbf{bundle together three different things}:
\begin{enumerate}
  \item Constructing an implicit ordinal outcome through
    hierarchical composition (complete tie-breaking/prioritized outcomes)
  \item Targeting a nonparametric effect measure defined via the
    ``pairs'' device (win probability/win ratio/win odds/etc.)
  \item Estimating that effect measure in a nonparametric/flexible
    manner using pairs, often also with devices for accounting for censoring
\end{enumerate}

This paper is focused on the first of these steps only.

Why is it useful to consider the three steps separately?

Consider some of the critiques of win methods in
\citet[e.g.~][]{ajufoFallaciesUsingWin2023}

\begin{itemize}
  \item Situating methods in their proper context can help
    trialists make informed choices among them
  \item Win methods can make use of
    plots and diagnostics for general ordinal outcomes, for the
    \emph{implicit} ordinal variable
    \begin{itemize}
      \item
        comparative CDF plots for the
      \item ridit plots
        \citep{brossHowUseRidit1958,jansenRiditAnalysisReview1984}
        relative to a reference group, which should allow one to
        display both full eCDF information and also win statistics
        (based on the interpretation of the mean ridit as a win probability)
    \end{itemize}e.g.
  \item Ordinal regression models can use win methods for model
    checking; it is
    also possible to present a model-based win effect measure
    alongside a common
    odds ratio for example. See
    \citep{agrestiOrdinalProbabilityEffect2017} for example, which
    covers both latent variable and predicted probability versions.
  \item The outcomes \emph{implicitly} constructed using
    DOOR and win methods may contain many levels, and may contain
    levels where an
    effect is not so meaningful (e.g. time-to-death when the followup time is
    short). By \emph{explicitly} constructing the associated
    ordinal outcome one
    might be forced to be more aware of this
  \item What is a reasonable win
    analysis of a specific component score? From the traditional
    ordinal endpoint
    school of thought what would make most sense is to \emph{only
      tie-break for the
    lowest and highest levels of that score}, and e.g. plot eCDFs of this new
    variable. Is this what the plots do that break down wins and losses by
    component score?
\end{itemize}

Disadvantages of generalized pairwise comparison approaches
\begin{itemize}
  \item Inherited from downsides of complete tiebreaking
  \item Inherited from downsides of nonparametric/pairwise effect measures
  \item Regression methods exist, but are newer and less battle-tested
\end{itemize}

\section{Reasons to combine ordinal outcomes}

\subsection{Providing a more holistic measure of benefit and harm}

\subsection{Incorporating intercurrent events, especially death,
for validity}

The intercurrent event could be (1) incorporated as a level (2) used
for tie-breaking or overriding.

This is a version of the ``composite strategy'' as described in e.g.
\citet{kahanEstimandsFrameworkPrimer2024}, but we expect it to yield
more efficiency than creating a \emph{binary} composite outcome.

For example overriding a numeric outcome by death.
Example: REMAP-CAP: organ-free survival, overridden by death
(Figure \ref{fig:remap_cap_override}).

\begin{figure}
  \includegraphics[width=5in]{remap_outcome_extra_cropped.jpg}
  \caption{Ordinal outcome used in the REMAP-CAP adaptive
    platform trial
    \citep{theremap-capinvestigatorsInterleukin6ReceptorAntagonists2021}.
    Illustrates overriding a numeric ordinal outcome (organ
    support-free days) by a binary ordinal outcome (death; this is
    ``sequential composition''), which
    would also be an intercurrent event. This is another example of
  sequential composition.}
  \label{fig:remap_cap_override}
\end{figure}

\subsection{Increasing precision by reducing the probability of the
largest category}

Theory of \citep{whiteheadSampleSizeCalculations1993} says that the relative
efficiency (variance) of an unadjusted treatment comparison using an ordinal
versus continuous outcome in the proportional odds model is given by $1 - \sum
\overline p_c^3$ where $\overline p_c$ is the marginal probability (averaged
over control and experimental groups) of outcome category $c$.

Same should apply to the Mann-Whitney-Wilcoxon test, and therefore to
the win probability, and therefore at least approximately to win
difference/win ratio/win odds.

Efficiency is maximized at $1 - 1/C^2$ where $C$ is the total number of
categories when the probabilities are all equal: $\overline p_c = 1/C$.

We have simulation results and some maths to show that the value of $\sum_c
\overline p_c^3$ is \emph{essentially a function of the largest category
probability} $\overline{p_{\max}} = \max \overline{p_c}$. See Figure
\ref{fig:p_max} for an illustration of the maths.

\begin{figure}
  \includegraphics[width=6in]{p_max_controls_efficiency.png}
  \caption{Bounds on the relative efficiency of an ordinal outcome
    compared to its continuous counterpart as a function of the maximum
    category probability $p_\text{max}$. The upper bound (best case
    relative efficiency for a given $p_\text{max}$) is always achieved in the
    limit of infinitely many categories ($C = \infty$). The lower bound
    is always achieved when the number of categories is as small as
  possible for a given $p_\text{max}$.}
  \label{fig:p_max}
\end{figure}

Thus simple advice is: for statistical efficiency, try to make the proportion
of patients in the largest category small.

However a converse is: if the proportion of patients in the largest
category is
small, not much is gained by tie-breaking.

\subsection{Augmenting a state outcome with longitudinal
information}

Distinguish between ``state'' and ``trajectory'' variables. State
variables summarize patient outcomes at a particular moment in time.
Trajectory variables, including times-to-event, summarize patient
state over a timecourse.

Trajectory and state variables can be incorporated into an ordinal outcome. For
example (from PASSPORT) ``worst pSOFA while hospitalized until day 28'' and
``length of hospitalization'', which are longitudinal variables (summary of a
trajectory), are combined with the day 28 state variable ``death''.

Incorporating trajectory variables into an ordinal outcome may seem unnatural.
But arguably most time-to-event analyses are in fact analyses of an
ordinal outcome. The Cox model for time-to-event data treats time as ordinal:
because (1) the baseline hazard is not modelled, (2) the model itself is
invariant to monotone rescalings of the time axis, and (3) the (partial)
likelihood (ignoring ties) depends only on the ranks of the data (i.e., is a
marginal likelihood). The Cox model is essentially an ordinal
stopping ratio model with complementary log-log link. (Perhaps a
difference is censoring.)

\section{Reasons to coarsen ordinal outcomes}

\subsection{Coarsening to avoid directly comparing levels}

In the PASSPORT outcome (Figure \ref{fig:passport_outcome}), for
example, we include both digit
amputation and ECMO
any time within 28 days post-enrollment under the heading of
``adverse clinical events''.

Including these outcomes within the same level of an
ordinal outcome does not mean assuming that the two outcomes have the
\emph{same} utility. Rather it represents \emph{avoiding comparing}
their utilities. (This was something clinicians asked about more than once.)

Also, we "chunk" longer hospital stays more aggressively to avoid
small categories, for interpretability's sake.

\subsection{Coarsening for ease of communication}

Too many levels can make it hard to directly compare meaningful
proportions/probabilities.

(It isn't really a problem for model fitting with modern software.)

Coarsening, including combining levels and replacing complete
tie-breaks with overrides, could help with this.

\section{Conclusions}

\newpage

\printbibliography

\end{document}